---
title: "QSIDE - Q1"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2024-03-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions for Generating QSIDE Data Crunching App Figures

### 1. Install necessary packages in R. These include ggplot2, tidyverse, and dplyr. To do so, use install.packages().

### 2. Inspect the sample dataset. For the R Shiny app, the QSIDE team supplied Data from North Carolina.

```{r}
policingdata=read.csv("durhamdata.csv")
head(policingdata)
```
### Format the DataFrame using tidyverse/dplyr. The steps are described below. 

### First, check if the necessary data is available in the DataFrame. For this analysis, we will assume that the data is all available since we are using a sample dataset. To check if data is available in the running Shiny App, we use an if statement that looks like the following : if (all(c("Race", "Officer") %in% colnames(policingdata))

### For the Q9, the race and officer variables must be available to answer the question. If there is missing data in these columns, filter them out. 

### For this situation, we need to first find the officers with the most incident reports using summarise/count in dplyr. We then select data from these officers and create subcategories for different racial groups. 

### Question: What proportion of police incidents go to different race groups for each of the most active officers?

```{r}
officertally <- policingdata %>%
  group_by(Officer) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

top25percent <- officertally %>%
  filter(count >= quantile(count, 0.75)) %>%
  pull(Officer)

top10 <- officertally %>%
  .[1:10, ] %>%
  pull(Officer)  

officerlist <- intersect(top25percent, top10)

qdata <- policingdata %>%
  filter(Officer %in% officerlist) %>%
  mutate(Officer = droplevels(as.factor(Officer))) %>%
  group_by(Officer, Race) %>%
  summarise(count = n()) %>%
  ungroup %>%
  complete(Officer, Race, fill = list(count = 0)) %>%
  group_by(Officer) %>%
  mutate(proportion = prop.table(count)) %>%
  filter(Race != "Missing race data") 

```

### From there, you can run plots in R using the filtered qdata DataFrame. To display this data, we are using a horizontal bar graph. 

```{r}
p <- ggplot() +
  geom_col(data = qdata, aes(x = Officer, y = proportion), fill="skyblue4") +
  ylab("Proportion") +
  scale_x_discrete(label=function(x) abbreviate(x, minlength=5))+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1), text = element_text(size = 15))
  p + facet_wrap(~Race)
```

















