---
title: "QSIDE - Q13"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2024-03-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions for Generating QSIDE Data Crunching App Figures

### 1. Install necessary packages in R. These include ggplot2, tidyverse, and dplyr. To do so, use install.packages().

### 2. Inspect the sample dataset. For the R Shiny app, the QSIDE team supplied Data from North Carolina.

```{r}
policingdata=read.csv("durhamdata.csv")
head(policingdata)
```
### Format the DataFrame using tidyverse/dplyr. The steps are described below. 

### First, check if the necessary data is available in the DataFrame. For this analysis, we will assume that the data is all available since we are using a sample dataset. To check if data is available in the running Shiny App, we use an if statement that looks like the following : if (all(c("Race", "Gender", "SearchConducted") %in% colnames(policingdata))

### For the Q13, the race, search conducted, and gender variables must be available to answer the question. If there is missing data in these columns, filter them out. 

### For this situation, we need to calculate mean bond amount using summarise in dplyr. 

```{r}
qdata <- policingdata %>%
  filter(SearchConducted == "Yes") %>%
  group_by(Race, Gender) %>%
  summarise(incidents = n(), .groups = 'drop') %>%
  ungroup() %>%
  complete(Race, Gender, fill = list(incidents = 0)) %>%
  filter(Race != "Missing race data" & Gender != "Missing gender data")
```

### From there, you can run plots in R using the filtered qdata DataFrame. To display this data, we are using a horizontal bar graph. 

```{r}
p <- qdata %>%
  ggplot(aes(x = incidents, y = Race)) +
  geom_col(aes(fill=Gender), position = "dodge", width=0.8) +
  ylab("Race") +
  xlab("Number of Searches") +
  scale_fill_manual(values = c("tomato3", "skyblue4")) +
  theme_bw() +
  theme(legend.position = "bottom", text = element_text(size = 15))
p
```

















