---
title: "QSIDE - Q6"
output: html_document
date: "2024-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions for Generating QSIDE Data Crunching App Figures

### 1. Install necessary packages in R. These include ggplot2, tidyverse, and dplyr. To do so, use install.packages().

### 2. Inspect the sample dataset. For the R Shiny app, the QSIDE team supplied Data from North Carolina.

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
```

```{r}
acsracegender = read.csv("acsracegender.csv")
head(acsracegender)
policingdata = read.csv("policingdata.csv")
head(policingdata)
```

### Format the DataFrame using tidyverse/dplyr. The steps are described below. 

### First, check if the necessary data is available in the DataFrame. For this analysis, we will assume that the data is all available since we are using a sample dataset. To check if data is available in the running Shiny App, we use an if statement that looks like the following : if (all(c("Race", "Gender", "Arrest") %in% colnames(policingdata))) {






### For the Q6, the race, gender, and arrest variables must be available to answer the question. If there is missing data in these columns, filter them out. 

### For this situation, we want to answer the question: What proportion of incidents result in arrest for different gender/race?

```{r}
qdata_policing <- policingdata %>%
  group_by(Race, Gender) %>%
  summarise(arrests = sum(Arrest, na.rm = TRUE), incidents = n(), .groups = 'drop') %>%
  ungroup() %>%
  complete(Race, Gender, fill = list(arrests = 0, incidents = 0)) %>%
  mutate(proportion = arrests / (incidents + .Machine$double.eps)) %>%
  select(-arrests, -incidents) %>%
  filter(Gender != "Missing gender data") %>%  
  filter(Race != "Missing race data") %>%
  mutate(datatype = "Policing Records")

qdata_population <- acsracegender %>% 
  group_by(Race, Gender) %>% 
  mutate(datatype = "Local Population") %>%
  ungroup() %>%
  arrange(Race, Gender)

qdata_combined <- left_join(qdata_policing, qdata_population, by = c("Race", "Gender")) %>%
  mutate(proportion = proportion.x * proportion.y) %>%
  select(Race, Gender, proportion = proportion, datatype.x) %>%
  rename(datatype = datatype.x)

qdata <- bind_rows(qdata_combined, expected %>% select(Race, Gender, proportion, datatype))

qdata <- qdata %>% 
  complete(Race, Gender, datatype, fill = list(proportion = 0)) %>%
  mutate(datatype = factor(datatype)) %>%
  mutate(datatype = relevel(datatype, ref = "Policing Records")) %>% 
  filter(Race != "Missing race data") %>% 
  filter(Gender != "Missing gender data")
```

### From there, you can run plots in R using the filtered qdata DataFrame. 

```{r}
 p <- qdata %>%
            ggplot(aes(x = Race, y = proportion, fill = Gender, alpha = datatype)) +
            geom_bar(stat = "identity", position = "stack") + 
            scale_fill_manual(values = c("coral", "blue"),
                              guide = guide_legend(title = "Gender", order = 1), 
                              labels = c("Man", "Women")) + 
            scale_alpha_manual(values = c(0.3, 0.7),
                               guide = guide_legend(title = "Data Type", order = 2), 
                               labels = c("Policing Records", "Local Population")) +
            facet_wrap(~Gender)+ 
            theme_bw() +
            theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
p
```
