---
title: "QSIDE -  Q5"
output: html_document
date: "2024-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions for Generating QSIDE Data Crunching App Figures

### 1. Install necessary packages in R. These include ggplot2, tidyverse, and dplyr. To do so, use install.packages().

### 2. Inspect the sample dataset. For the R Shiny app, the QSIDE team supplied Data from North Carolina.

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
```

```{r}
acsracegender = read.csv("acsracegender.csv")
head(acsracegender)
policingdata = read.csv("policingdata.csv")
head(policingdata)
```

### Format the DataFrame using tidyverse/dplyr. The steps are described below. 

### First, check if the necessary data is available in the DataFrame. For this analysis, we will assume that the data is all available since we are using a sample dataset. To check if data is available in the running Shiny App, we use an if statement that looks like the following : if (all(c("Race", "Gender", "Arrest") %in% colnames(policingdata))) {





### For the Q5, the race, gender, and arrest variables must be available to answer the question. If there is missing data in these columns, filter them out. 

### For this situation, we want to answer the question: What proportion of incidents related to quality of life violations go to different gender/race groups?

```{r}
observed <- policingdata %>%
            filter(Arrest == TRUE) %>%
            group_by(Race, Gender) %>%
            summarise(people = n()) %>%
            ungroup %>%
            complete(Race, Gender, fill = list(people = 0)) %>%
            filter(Race != "Missing race data") %>%
            filter(Gender != "Missing gender data") %>%
            arrange(Race, Gender)
          
expected <- acsracegender %>%
            complete(Race, Gender, datatype, fill = list(proportion = 0)) %>%
            filter(Race != "Missing race data") %>%
            filter(Gender != "Missing gender data") %>%
            arrange(Race, Gender)

expected <- expected %>%
  mutate(datatype = "Local Population")

qdata <- observed %>%
            mutate(proportion = prop.table(people)) %>%
            select(-people) %>%
            mutate(datatype = "Policing Records") %>%
            bind_rows(expected %>% select(Race, Gender, proportion, datatype)) %>%
            complete(Race, Gender, datatype, fill = list(proportion = 0)) %>%
            mutate(datatype = factor(datatype)) %>%
            mutate(datatype = relevel(datatype, ref = "Policing Records")) %>%
            filter(Gender != "Missing gender data") %>%
            filter(Race != "Missing race data")
```

### From there, you can run plots in R using the filtered qdata DataFrame. 

```{r}
 p <- qdata %>%
            ggplot(aes(x = Race, y = proportion, fill = Gender, alpha = datatype)) +
            geom_bar(stat = "identity", position = "stack") +
            xlab("Race") + 
            ylab("Proportion") + 
            scale_fill_manual(values = c("coral", "blue"),
                              guide = guide_legend(title = "Gender"), 
                              labels = c("Man", "Women")) + 
            scale_alpha_manual(values = c(0.3, 0.7),
                               guide = guide_legend(title = "Data Type"),
                               labels = c("Policing Records", "Local Population")) +
            facet_wrap(~Gender)+ 
            theme_bw() +
            theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
p
```